---
title: "Emissions v Mass from EEA"
output: html_notebook
---

```{r}
knitr::opts_chunk$set(results='hide', message=FALSE, warning=FALSE)
```

Load libraries...
```{r, echo=F}
library(httr)
library(tidyjson)
library(dplyr) # mutating etc.
library(stringr) # manipulating e.g. manufacturer names
library(tidyr) # for lm fits to go with faceted graphs
library(purrr) # for lm fits to go with faceted graphs
library(broom) # for lm fits to go with faceted graphs
library(plotly) # for zoomable/interactive graphics; includes ggplot
```

Function to get EEA data...
```{r, results='hide', message=FALSE, warning=FALSE}
eeaGet <- function(query){
  tryCatch(return({
    GET("https://discodata.eea.europa.eu/sql",
                        query = list(query=query))%>%
    content(as="parsed") %>%
    (function(x){x$results}) %>%
    spread_all
  }), error = function(e){print(paste("problem retrieving",query))})
}
```

Functions to clean up the harmonized manufacturer names...
```{r, results='hide', message=FALSE, warning=FALSE}
eeaCleanNames <- function(df){ # just a function to rewrite some of these names
  df %>%
    mutate(name = toupper(Mh)) %>%
    mutate(name = str_replace(name, "GENERAL MOTORS?", "GM")) %>%
    mutate(name = str_replace(name, "LAND ROVER","LAND-ROVER")) %>%
    mutate(name = str_replace(name, "-WERKE|-AMG","")) %>%
    mutate(name = str_replace(name, "ROLLS ROYCE","ROLLS-ROYCE")) %>%
    mutate(name = str_replace(name, "VEHICULES ELECTRIQUES PININFARINA-BOLLORE","PININFARINA-BOLLORE")) %>% 
    mutate(name = str_replace(name, "UNKNWON","UNKNOWN")) %>%
    mutate(name = str_replace(name, "GREAT WALL","GREAT-WALL")) %>%
    mutate(name = str_replace(name, "SHANGHAI MAPLE","MAPLE")) %>%
    mutate(name = str_replace(name, "^ROVER","TATA")) %>%
    mutate(name = str_replace(name, "AUTOMOBILES?\\s?","")) %>%
    mutate(name = str_replace(name, "MOTOR CORPORATION\\s?","")) %>%
    mutate(name = str_replace(name, "(GENERAL )MOTOR COMPANY\\s?","")) %>%
    mutate(name = str_replace(name, "MOTORS?\\s?","")) %>%
    mutate(name = if_else(str_detect(Mh, "SHUANGHUAN"), "SHUANGHUAN", name)) %>%
    mutate(name = if_else(str_detect(Mh, "BAYERISCHE MOTOREN WERKE AG"), "BMW", name)) %>%
    mutate(first = str_extract(name, "(^[^\\s]*)(\\s|$)")) %>%
    mutate(first = str_squish(str_trim(first))) %>%
    mutate(man.name.clean = first, first = NULL, name = NULL) 
}

eeaRegroupByCleanNames <- function(df){
  df %>% # renamed cleanly
    group_by(man.name.clean, year) %>%
    summarise(mass_avg = weighted.mean(mass_avg, n_sales), # and then re-group and re-summarize according to the new names
              wltp_avg = weighted.mean(wltp_avg, n_sales),
              enedc_avg = weighted.mean(enedc_avg, n_sales),
              n_sales = sum(n_sales)) %>%
    ungroup
}

# looking at specific odd names
# mass.tbl %>%
#   filter(str_detect(Mh, "LIFAN"))
```


Functions to make the bubble graphs...
```{r, results='hide', message=FALSE, warning=FALSE}
eeaBubbleGraph <- function(df, x, y, title){ # we'd like to make bubble size a var as well, but it's complicated!

    # first, the fit to be able to label them in the graphs: 
  df %>%
    nest(-year) %>% # fold everything by year
    filter(year>2016) %>% # use only years after 2016, because no emissions data before then
    mutate(fit = map(data, ~lm(paste(y,"~",x), data = ., weights = n_sales)),
           results = map(fit, tidy)) %>% # fit and clean the fit for each year
    unnest(results) %>% # unfold the fit results
    pivot_wider(id_cols=c(year, data), values_from = estimate, names_from = term) %>% # make a column for each fit result of interest
    rename(slope=mass_avg, intercept=`(Intercept)`) %>% # 
    mutate(x_max = map(data, ~(max(.$mass_avg)))) %>% 
    unnest(x_max) %>%
    mutate(y_max = x_max*slope + intercept, x_max = x_max-100, man.name.clean = "name", mass_avg = 1, wltp_avg = 1, n_sales = 1) ->
    emissions.ratio
  
  # then, make the actual plots:
  fig <- (ggplot(filter(mass.tbl.clean, year>2016), aes_string(label="man.name.clean", x=x, y=y, size="n_sales")) +
    facet_grid(cols = vars(year),drop = T) +
    geom_point(alpha=0.5) +
    ylab("Specific emissions fleet average (g/km)") +
    xlab("Average vehicle mass (kg)") + 
    labs(size = "Sales by manufacturer") +
    ggtitle(title) +
    geom_smooth(method = "lm", mapping = aes(weight = n_sales)) +
    #stat_smooth(aes(outfit=fit<<-..y..))+
    geom_text(data=emissions.ratio, aes(x=x_max, y=y_max, label=round(slope, digits = 3)), size=5, nudge_x = -100)) %>%
    ggplotly %>%
    layout(automargin=T) %>%
    show
}
```

### Getting the data

"Mh" seems to have the fewest missing values, and also the fewest distinct categories, so lets use that one.
We'll fetch _average mass_ and _average emissions_ by _year_ and _"harmonized manufacturer name" (Mh)_.


Get aggregate European data...
```{r, results='hide', message=FALSE, warning=FALSE}
mass.tbl <- eeaGet("SELECT year, Mh, 
                                          AVG(CAST([m (kg)] AS FLOAT)) mass_avg, 
                                          AVG(CAST([Ewltp (g/km)] AS FLOAT)) wltp_avg,
                                          AVG(CAST([Enedc (g/km)] AS FLOAT)) enedc_avg,
                                          SUM(r) n_sales
                      FROM [CO2Emission].[latest].[co2cars]
                      GROUP BY year, Mh")
```

Clean up the European data...
```{r, results='hide', message=FALSE, warning=FALSE}
mass.tbl %>% 
  eeaCleanNames %>%
  eeaRegroupByCleanNames->
mass.tbl.clean
```

Get data just for Germany, nicely grouped...
```{r, results='hide', message=FALSE, warning=FALSE}
mass.de.tbl <- eeaGet("SELECT year, Mh,
AVG(CAST([m (kg)] AS FLOAT)) mass_avg, 
AVG(CAST([Ewltp (g/km)] AS FLOAT)) wltp_avg,
AVG(CAST([Enedc (g/km)] AS FLOAT)) enedc_avg,
SUM(r) n_sales
FROM [CO2Emission].[latest].[co2cars]
WHERE MS='DE'
GROUP BY year, Mh")
```

Clean up the German data...
```{r, results='hide', message=FALSE, warning=FALSE}
mass.de.tbl %>%
  eeaCleanNames %>%
  eeaRegroupByCleanNames ->
  mass.de.tbl.clean
```

## Bubble graphs with weighted fits

### Graphs for Europe

First, WLTP vs. mass:
```{r, results='hide', message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
eeaBubbleGraph(mass.tbl.clean, "mass_avg", "wltp_avg", "WLTP emissions v. mass (EU manufacturer fleet averages)")
```
Second, NEDC vs. mass:
```{r, results='hide', message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
eeaBubbleGraph(mass.tbl.clean, "mass_avg", "enedc_avg", "NEDC emissions v. mass (EU manufacturer fleet averages)")
```

### Graphs for Germany
First, WLTP vs. mass:
```{r, results='hide', message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
eeaBubbleGraph(mass.de.tbl.clean, "mass_avg", "wltp_avg", "WLTP emissions v. mass (DE manufacturer fleet averages)")
```
Second, NEDC vs. mass:
```{r, results='hide', message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
eeaBubbleGraph(mass.de.tbl.clean, "mass_avg", "enedc_avg", "NEDC emissions v. mass (DE manufacturer fleet averages)")
```