---
title: "EEA data: Flottengrenzwerte"
output: html_notebook
---
  

```{r}
library(httr)
library(tidyjson)
library(readr)
library(dplyr)
library(data.table)
library(ggrepel) # for hopefully better labels
library(plotly) # for zoomable/interactive graphics
library(broom) # for lm fits to go with faceted graphs
```

## Test getting the EEA stuff
```{r}
tbl <- GET("https://discodata.eea.europa.eu/sql",
                      query = list(query="SELECT year, Mh, 
                                          [m (kg)] AS mass, 
                                          [Ewltp (g/km)] AS wltp,
                                          [Enedc (g/km)] AS enedc,
                                          r AS n_sales
                      FROM [CO2Emission].[latest].[co2cars]"))%>%
tbl %>%  content(as="parsed") %>%
  (function(x){x$results}) %>%
  spread_all 

```


## First, take a look at some of these variables.

What are their proper names?
```{r}
eeaGet <- function(query){
  tryCatch(return({
    GET("https://discodata.eea.europa.eu/sql",
                        query = list(query=query))%>%
    content(as="parsed") %>%
    (function(x){x$results}) %>%
    spread_all
  }), error = function(e){print(paste("problem retrieving",query))})
}

all.vars <- names(eeaGet("SELECT TOP 1 * FROM [CO2Emission].[latest].[co2cars]"))

all.vars
```
What are the distinct categories in each variable we're looking at?
```{r}
queryDistinct <- function(var){
  query = paste0("SELECT DISTINCT [", 
                                                var,
                                                "] FROM [CO2Emission].[latest].[co2cars]")
  print(query)
  tryCatch(return (eeaGet(query) %>% pull(var)), error = function(e){print(paste("problem retrieving",var))})
}

vars <- c("Ft", "Fm", "Mf (kg)", "Cr", "Mh", "Man", "Mk")

sapply(vars, queryDistinct)

```
## Okay! Get average mass [m (kg)] and emissions [Ewltp (g/km), Enedc (g/km)] and sum of vehicles sold [r] for each fuel type [Ft] for each year [year]

```{r}
mass.emissions.by.ft <- GET("https://discodata.eea.europa.eu/sql",
                      query = list(query="SELECT year, Ft, 
                                          AVG(CAST([m (kg)] AS FLOAT)) mass_avg, 
                                          SUM(CAST([m (kg)] AS FLOAT)) mass_tot,
                                          AVG(CAST([Ewltp (g/km)] AS FLOAT)) wltp_avg,
                                          SUM(CAST([Ewltp (g/km)] AS FLOAT)) wltp_tot,
                                          AVG(CAST([Enedc (g/km)] AS FLOAT)) enedc_avg,
                                          SUM(CAST([Enedc (g/km)] AS FLOAT)) enedc_tot,
                                          SUM(r) n_total
                      FROM [CO2Emission].[latest].[co2cars]
                      GROUP BY year, Ft"))%>%
  content(as="parsed") %>%
  (function(x){x$results}) %>%
  spread_all %>%
  mutate(Ft=tolower(Ft))

```

Try a little viz--over time by category
```{r}
mass.emissions.by.ft %>% 
  mutate(Ft=tolower(Ft)) %>%
  group_by(Ft,year) %>%
  summarize(mass_tot=sum(mass_tot,na.rm = TRUE),
            wltp_tot=sum(wltp_tot,na.rm = TRUE),
            enedc_tot=sum(enedc_tot,na.rm = TRUE),
            n_tot=sum(n_total,na.rm = TRUE)) ->
  summarized

summarized %>% distinct(Ft,year)
```
 try better palette...
```{r}
#install.packages("pals")
library(pals)
```
 

```{r}
library(ggplot2)
# total mass
ggplot(mass.emissions.by.ft, aes(x=year, y=mass_tot)) +
  geom_point(aes(color=Ft)) + 
  scale_fill_gradientn(colours = stepped())
```
average mass...
```{r}
ggplot(mass.emissions.by.ft, aes(x=year, mass_avg)) +
  geom_point(aes(color=Ft))
```

average emissions...
```{r}
# WLTP
ggplot(mass.emissions.by.ft, aes(x=year, y=wltp_avg)) +
  geom_point(aes(color=Ft))
```
```{r}

# NEDC
ggplot(mass.emissions.by.ft, aes(x=year, y=enedc_avg)) +
  geom_point(aes(color=Ft)) + 
  scale_color_discrete()
```
It seems like one of the biggest issues is the different names for same category. Let's fix that.
```{r}
mass.emissions.by.ft %>% distinct(Ft) %>% pull(Ft) 
```
```{r}

mass.emissions.by.ft %>%
  mutate(Ft = recode(na_if(Ft,""), 
                     "ng-biomethane" = "rng",
                     "ng_biomethane" = "rng",
                     "diesel-electric" = "hybrid_diesel",
                     "diesel/electric" = "hybrid_diesel",
                     "hybrid/petrol/e" = "hybrid_petrol",
                     .missing = "unknown",
                     "petrol-electric" = "hybrid_petrol",
                     "petrol-gas" = "petrol_gas",
                     "petrol phev" = "p_hybrid_petrol",
                     "petrol/electric" = "hybrid_petrol",
                     "NA" = "unknown"
                     )) %>%
  ggplot(aes(x=year, y=mass_avg)) +
  geom_point(aes(color=Ft)) +
  scale_color_manual(values=glasbey())
```
## Now, try bubble plot of mass and emissions over time by manufacturer.
First, check which of these manufacturer categories is most reasonable
```{r}
man.vars <- c("man","mh","Mk","MMS","MP")

man.vars.cat <- sapply(man.vars, queryDistinct)
sapply(man.vars.cat, length)
man.vars.cat$MP

mans <- lapply(man.vars,
               function(var){
                 print(var)
                 df <- eeaGet(paste0("SELECT ", var, ", COUNT(ID) n 
       FROM [CO2Emission].[latest].[co2cars]
       GROUP BY ", var))
                 print(names(df))
                 return(df)
               })

lapply(mans, function(df){
  dt <- as.data.table(df)
  print(names(dt)[2])
  print(paste("Missing:",dt[is.na(dt[[2]])|dt[[2]]==""|dt[[2]]=="-"|dt[[2]]=="UNKNOWN", sum(n)]))
  print(paste("Categories:", nrow(dt)))
  return (NULL)
})


mans
```
"Mh" seems to have the fewest missing values, and also the fewest distinct categories, so lets use that one.
get the appropriate data--average mass and emissions by year and manufacturer:
```{r}
mass.tbl <- eeaGet("SELECT year, Mh, 
                                          AVG(CAST([m (kg)] AS FLOAT)) mass_avg, 
                                          AVG(CAST([Ewltp (g/km)] AS FLOAT)) wltp_avg,
                                          AVG(CAST([Enedc (g/km)] AS FLOAT)) enedc_avg,
                                          SUM(r) n_sales
                      FROM [CO2Emission].[latest].[co2cars]
                      GROUP BY year, Mh")
# get rid of the JSON column--I don't know a better way :(
# mass.tbl <- as.tibble(as.data.table(mass.df))
```


How are our manufacturers looking, though?
```{r}
# write it out if you need to--this will require some conversion away from the json format
# mass.tbl %>% 
#   filter(year==2019) %>%
#   select(Mh, n_sales) %>%
#   write_csv("/Users/maita.schade/Downloads/manufacturers.csv")

mass.tbl %>% filter(Mh=="ISUZU") 
```
### We need to clean the manufacturers
I'm going to give OpenRefine a try here.
We have to the the same variables, and regroup based on sales numbers
```{r}
# OpenRefine is a mess--stopped doing this
# mass.dt <- as.data.table(mass.tbl)
# mass.dt[,("..JSON"):=NULL]
# fwrite(mass.dt, "/Users/maita.schade/Documents/Work/Flottengrenzwerte/out/eea_mass_emis.csv")
```

Seems like openrefine isn't very amenable to matching from beginning. I'll try to cluster these manufacturers.
```{r}
# clustering--also not very helpful!
# install.packages("stringdist")
# library(stringdist)
# uniqueman <- unique(as.character(mass.dt$Mh))
# distanceman <- stringdistmatrix(uniqueman,uniqueman,method = "jw")
# rownames(distanceman) <- uniqueman
# hc <- hclust(as.dist(distanceman))
# plot(hc)
# rect.hclust(hc,k=20)

eeaCleanNames <- function(df){ # just a function to rewrite some of these names
  df %>%
    mutate(name = toupper(Mh)) %>%
    mutate(name = str_replace(name, "GENERAL MOTORS?", "GM")) %>%
    mutate(name = str_replace(name, "LAND ROVER","LAND-ROVER")) %>%
    mutate(name = str_replace(name, "-WERKE|-AMG","")) %>%
    mutate(name = str_replace(name, "ROLLS ROYCE","ROLLS-ROYCE")) %>%
    mutate(name = str_replace(name, "VEHICULES ELECTRIQUES PININFARINA-BOLLORE","PININFARINA-BOLLORE")) %>% 
    mutate(name = str_replace(name, "UNKNWON","UNKNOWN")) %>%
    mutate(name = str_replace(name, "GREAT WALL","GREAT-WALL")) %>%
    mutate(name = str_replace(name, "SHANGHAI MAPLE","MAPLE")) %>%
    mutate(name = str_replace(name, "^ROVER","TATA")) %>%
    mutate(name = str_replace(name, "AUTOMOBILES?\\s?","")) %>%
    mutate(name = str_replace(name, "MOTOR CORPORATION\\s?","")) %>%
    mutate(name = str_replace(name, "(GENERAL )MOTOR COMPANY\\s?","")) %>%
    mutate(name = str_replace(name, "MOTORS?\\s?","")) %>%
    mutate(name = if_else(str_detect(Mh, "SHUANGHUAN"), "SHUANGHUAN", name)) %>%
    mutate(name = if_else(str_detect(Mh, "BAYERISCHE MOTOREN WERKE AG"), "BMW", name)) %>%
    mutate(first = str_extract(name, "(^[^\\s]*)(\\s|$)")) %>%
    mutate(first = str_squish(str_trim(first))) %>%
    mutate(man.name.clean = first, first = NULL, name = NULL) 
}

eeaRegroupByCleanNames <- function(df){
  df %>% # renamed cleanly
    group_by(man.name.clean, year) %>%
    summarise(mass_avg = weighted.mean(mass_avg, n_sales), # and then re-group and re-summarize according to the new names
              wltp_avg = weighted.mean(wltp_avg, n_sales),
              enedc_avg = weighted.mean(enedc_avg, n_sales),
              n_sales = sum(n_sales)) %>%
    ungroup
}
# looking at specific odd names
# man.names %>%
#   filter(str_detect(Mh, "LIFAN"))

mass.tbl %>% 
  eeaCleanNames %>%
  eeaRegroupByCleanNames->
mass.tbl.clean

# looking at specific odd cases
# filter(mass.tbl, str_starts(Mh, "ALPINA"))

mass.tbl.clean %>%
  group_by(man.name.clean) %>%
  summarise(total = sum(n_sales))

```
### Graphics for Europe

```{r}
# just for 2017... what is going on with the fits?!!
# All is well now, lm and geom_smooth match up perfectly. I was so confused :'-D
# mass.tbl.2017 <- filter(mass.tbl.clean, year==2017)
# lm.2017 <- lm("wltp_avg ~ mass_avg", data = mass.tbl.2017) #weights = n_sales, 
# mass.tbl.2017 %>%
#   mutate(lm.fit = predict.lm(lm.2017, .)) ->
#   mass.tbl.2017.fit
# 
# p <- ggplot(mass.tbl.2017.fit) +
#   geom_point(alpha=0.5, aes(x=mass_avg, y=wltp_avg, size=n_sales)) +
#   geom_line(aes(x=mass_avg, y=lm.fit), color="red") +
#   stat_smooth(method = "lm", mapping = aes(x=mass_avg, y=wltp_avg, weight = NULL), fullrange=T)
# p
```


Make a bubble graph!
```{r}
eeaBubbleGraph <- function(df, x, y, title){ # we'd like to make bubble size a var as well, but it's complicated!

  # first, the fit to be able to label them in the graphs: 
df %>%
  nest(-year) %>% # fold everything by year
  filter(year>2016) %>% # use only years after 2016, because no emissions data before then
  mutate(fit = map(data, ~lm(paste(y,"~",x), data = ., weights = n_sales)),
         results = map(fit, tidy)) %>% # fit and clean the fit for each year
  unnest(results) %>% # unfold the fit results
  pivot_wider(id_cols=c(year, data), values_from = estimate, names_from = term) %>% # make a column for each fit result of interest
  rename(slope=mass_avg, intercept=`(Intercept)`) %>% # 
  mutate(x_max = map(data, ~(max(.$mass_avg)))) %>% 
  unnest(x_max) %>%
  mutate(y_max = x_max*slope + intercept, x_max = x_max-100, man.name.clean = "name", mass_avg = 1, wltp_avg = 1, n_sales = 1) ->
  emissions.ratio

# then, make the actual plots:
(ggplot(filter(mass.tbl.clean, year>2016), aes_string(label="man.name.clean", x=x, y=y, size="n_sales")) +
  facet_grid(cols = vars(year),drop = T) +
  geom_point(alpha=0.5) +
  ylab("Specific emissions fleet average (g/km)") +
  xlab("Average vehicle mass (kg)") + 
  labs(size = "Sales by manufacturer") +
  ggtitle(title) +
  geom_smooth(method = "lm", mapping = aes(weight = n_sales)) +
  #stat_smooth(aes(outfit=fit<<-..y..))+
  geom_text(data=emissions.ratio, aes(x=x_max, y=y_max, label=round(slope, digits = 3)), size=5, nudge_x = -100)) %>%
  ggplotly
}
```

Graphs for Europe
```{r}
eeaBubbleGraph(mass.tbl.clean, "mass_avg", "wltp_avg", "WLTP emissions v. mass (EU manufacturer fleet averages)")
```
```{r}
eeaBubbleGraph(mass.tbl.clean, "mass_avg", "enedc_avg", "NEDC emissions v. mass (EU manufacturer fleet averages)")
```

### Germany only

Das gleiche, aber nur für Deutschland.
Wie sind die Länder ausgewiesen?

```{r}
eeaGet("SELECT DISTINCT(MS) FROM [CO2Emission].[latest].[co2cars]")
```

Get data just for Germany, nicely grouped
```{r}
mass.de.tbl <- eeaGet("SELECT year, Mh,
AVG(CAST([m (kg)] AS FLOAT)) mass_avg, 
AVG(CAST([Ewltp (g/km)] AS FLOAT)) wltp_avg,
AVG(CAST([Enedc (g/km)] AS FLOAT)) enedc_avg,
SUM(r) n_sales
FROM [CO2Emission].[latest].[co2cars]
WHERE MS='DE'
GROUP BY year, Mh")

mass.de.tbl %>%
  eeaCleanNames %>%
  eeaRegroupByCleanNames ->
  mass.de.tbl.clean
```

Make bubble graphs for Germany!
WLTP first...
```{r}
eeaBubbleGraph(mass.de.tbl.clean, "mass_avg", "wltp_avg", "WLTP emissions v. mass (DE manufacturer fleet averages)")
```
```{r}
eeaBubbleGraph(mass.de.tbl.clean, "mass_avg", "enedc_avg", "NEDC emissions v. mass (DE manufacturer fleet averages)")
```
```{r}
"Z (Wh/km)" %>% (function(var){
                 print(var)
                 df <- eeaGet(paste0("SELECT [", var, "], COUNT(ID) n 
       FROM [CO2Emission].[latest].[co2cars]
       GROUP BY [", var, "]"))
                 print(names(df))
                 return(df)
               }) %>%
(function(df){
  dt <- as.data.table(df)
  print(names(dt)[2])
  print(paste("Missing:",dt[is.na(dt[[2]])|dt[[2]]==""|dt[[2]]=="-"|dt[[2]]=="UNKNOWN", sum(n)]))
  print(paste("Categories:", nrow(dt)))
  return (NULL)
})

```

