---
title: "TCO interactive viz"
date: 11-11-2020
output: html_notebook
editor_options: 
  chunk_output_type: console
---
This is supposed to be a quick-and-dirty interactive graphic of TCO over time, with sliders to control key parameters
```{r}
library(ggplot2) # good plotting
library(dplyr) # manipulate tbls
library(readr) # read_csv
library(tidyr) # pivoting
library(manipulate) # interactive graphs
library(pals) # better color palettes
```

Load and clean up components from literature
```{r}
components.tbl <- read_csv('/Users/maita.schade/Documents/Work/Database collection/tco_cost_components.csv')
```
Currency
```{r}
usd.eur.conv.tbl <- tibble(year=c(2017,2018),usd.eur.conv=c(1.13,1.18))
components.tbl %>%
  left_join(usd.eur.conv.tbl, by=c("currency.year"="year")) %>%
  mutate(value = (if_else(is.na(value.eur), value.usd*usd.eur.conv, value.eur))) %>%
  select(-c(value.usd,value.eur,currency,usd.eur.conv)) ->
  components.tbl.eur
```

Other fixes
```{r}
components.tbl.eur %>%
  pivot_wider(names_from = cost.component, values_from = value) %>%
  mutate(deprec.tot = if_else(is.na(purchase.tot)|is.na(resale.tot), deprec.tot, purchase.tot-resale.tot)) %>%
  mutate(resale.tot = if_else(is.na(purchase.tot)|is.na(deprec.tot), resale.tot, purchase.tot-deprec.tot)) %>%
  mutate(annual.overhead = annual.overhead.tot / ownership.years, annual.overhead.tot = NULL) %>%
  pivot_longer(c(annual.overhead, battery, deprec.tot, driving.tot, financing, propellant, purchase.tot, resale.tot, subsidy.pos, variable.misc.tot, range.comp), 
                 values_drop_na = T, names_to = "cost.component") ->
  components.tbl.eur.cleaner
```

Analyze individual components
```{r,  fig.width=10,fig.height=10}
components.tbl.eur %>% distinct(cost.component)
components.tbl.eur.cleaner %>%
  ggplot(aes(x=value.year, y=value)) + 
    geom_point() +
    scale_color_manual(values=glasbey()) +
    facet_grid(cols = vars(engine.type), rows = vars(cost.component))
```

### inputs
km.yr #constant
p.eur.L # constant--1.20-1.70
d.eur.L # constant--1.00-1.50
e.eur.kwh # constant--
batt.eur.kwh.exp
kwh
subsidy
deprec.perc
margin
ownership.yrs
 
### model
```
purchase.tot.e_slider * yr * (1-exp(deprec.factor_slider*ownership.yrs_slider)) [SEPARATE slider for purchase price dev e,p,d]+ 
(subsidy.pos_slider) + 
(km.yr_slider * ownership.yrs_slider * e.eur.kwh_slider * kwh.km) [OR p/d equivalent] +
annual.overhead_slider*annual.overhead.factor * ownership.yrs

```

```{r}
x <- seq(1,100)
a <- 5
plot(x, 50000*exp(-0.3*x))
```

Get linear fits to some of these

```{r fig.width=10,fig.height=12}

components.tbl.eur.cleaner %>%
  ggplot(aes(x=value.year, y=value)) + 
    geom_point() +
    scale_color_manual(values=glasbey()) +
    facet_grid(cols = vars(engine.type), rows = vars(cost.component)) +
  geom_smooth(method = "lm")
```

```{r}
library(broom)
library(tidyverse)
components.tbl.eur.cleaner %>%
  nest(-engine.type, -cost.component) %>%
  mutate(fit = map(data, ~ lm("value ~ value.year", .)),
         results = map(fit, tidy)) %>%
  unnest(results) %>%
  select(engine.type, cost.component, term, estimate, std.error)

```
```{r}
library(broom)
library(tidyverse)
components.tbl.eur.cleaner %>%
  nest(-engine.type, -cost.component) %>%
  mutate(fit = map(data, ~ lm("value ~ 1", .)),
         results = map(fit, tidy)) %>%
  unnest(results) %>%
  select(engine.type, cost.component, term, estimate, std.error) %>%
  filter(cost.component=="annual.overhead") ->
  annual.overhead.estimates

annual.overhead.estimates %>%
  mutate(overhead.factor = estimate/filter(annual.overhead.estimates, engine.type=="bev")$estimate) %>%
  select(engine.type, overhead.factor)
```


```{r}
glance(lm(value ~ value.year, components.tbl.eur.cleaner))
```




```{r}
# $/kWh
cost.tbl <- tibble(year=c(2010:2030), cost=1000)
```
Test manipulate
```{r, R.options=list(chunk_output_type="console")}
# install.packages("manipulate")
library(manipulate)
# run this to deactivate manipulate bug
# manipulate(plot(1:5, cex=size), size = slider(0.5,10,step=0.5))
manipulate(
  {ggplot(cost.tbl, aes(year, 1000*exp(coeff*(year-2010)) - subv)) +
   geom_line() + ylim(0,1000)},
  coeff = slider(-0.3, 0),
  subv = slider(0, 100)
)
```
Build a manipulable table and plot
```
DEPRECIATION = (purchase.tot.e.0 - purchase.tot.e_slider * yr) * (1-exp(deprec.factor_slider*ownership.yrs_slider)) [SEPARATE slider for purchase price dev e,p,d]+ 
SUBSIDY = (subsidy.pos_slider) + 
DRIVING = (km.yr_slider * ownership.yrs_slider * e.eur.kwh_slider * kwh.km) [OR p/d equivalent] +
OTHER FIXED = annual.overhead_slider*annual.overhead.factor * ownership.yrs
```


```{r}
plotTableF <- function(purchase.tot.bev = 2000, 
                       purchase.tot.diesel = 1000, 
                       purchase.tot.petrol = 1000, 
                       deprec.factor = -0.3, 
                       ownership.yrs= 8, 
                       subsidy.pos = 4000, 
                       km.yr = 15000, 
                       e.eur.kwh = 0.3, 
                       petrol.eur.L = 1.5, 
                       diesel.eur.L = 1, 
                       kwh.km = 0.189, 
                       petrol.L.km = 0.08, 
                       diesel.L.km = 0.07,
                       annual.overhead = 810){
  plotting.tbl <- expand_grid(year = seq(2015, 2030), engine.type = c("bev","diesel","petrol"))
  plotting.tbl %>%
    mutate(Subsidy = if_else(engine.type=="bev", -subsidy.pos, NULL)) %>%
    mutate(Depreciation = if_else(engine.type=="bev", (50000-purchase.tot.bev*(year-2015) + Subsidy)*(1-exp(deprec.factor*ownership.yrs)),
           if_else(engine.type=="petrol", (38000-purchase.tot.petrol*(year-2015))*(1-exp(deprec.factor*ownership.yrs)),
                   if_else(engine.type=="diesel",(38000-purchase.tot.diesel*(year-2015))*(1-exp(deprec.factor*ownership.yrs)),NULL)
                   ))) %>%
    mutate(Driving = if_else(engine.type=="bev", km.yr * ownership.yrs * e.eur.kwh * kwh.km,
           if_else(engine.type=="petrol", km.yr * ownership.yrs * petrol.eur.L * petrol.L.km,
                   if_else(engine.type=="diesel", km.yr * ownership.yrs * diesel.eur.L * diesel.L.km, NULL)
                   ))) %>%
    mutate(Fixed = if_else(engine.type=="bev", annual.overhead * ownership.yrs,
           if_else(engine.type=="petrol", annual.overhead * 1.203294 * ownership.yrs,
                   if_else(engine.type=="diesel", annual.overhead * 1.044825 * ownership.yrs, NULL)
                   ))) %>%
   pivot_longer(c(Depreciation, Subsidy, Driving, Fixed), 
                 values_drop_na = T, names_to = "cost.component")
}

test.tbl <- plotTableF()

ggplot(test.tbl, aes(x = engine.type, y = value, fill=cost.component)) +
  geom_col(position = "stack") + 
  facet_grid(cols=vars(year)) +
  theme(axis.text.x = element_text(angle = 90))
```
```{r}
manipulate(
  {ggplot(cost.tbl, aes(year, 1000*exp(coeff*(year-2010)) - subv)) +
   geom_line() + ylim(0,1000)},
  coeff = slider(-0.3, 0),
  subv = slider(0, 100)
)
manipulate(
  {ggplot(plotTableF(purchase.tot.bev = purchase.tot.bev), aes(x = engine.type, y = value, fill=cost.component)) +
      geom_col(position = "stack")
      # facet_grid(cols=vars(year)) +
      # ylab("TCO") +
      # theme(axis.text.x = element_text(angle = 90))
    },
  purchase.tot.bev = slider(0,3000,initial = 2000) 
  
)
```
Schönere Farben...
Agora Farben
```{r}
agora.cols.prim.str <- c("29 / 74 / 98",
                         "0 / 72 / 153",
                         "0 / 106 / 171",
                         "123 / 192 / 188")
agora.cols.1.str <- c("162 / 140 / 171",
                      "50 / 33 / 64",
                      "62 / 163 / 220",
                      "42 / 106 / 178",
                      "244 / 221 / 197")
agora.cols.2.str <- c("138 / 170 / 164",
                      "230 / 231 / 226",
                      "214 / 217 / 190",
                      "239 / 228 / 141",
                      "169 / 157 / 110")
agora.cols.3.str <- c("65 / 61 / 73",
                      "45 / 118 / 102",
                      "86 / 162 / 77",
                      "124 / 167 / 171",
                      "207 / 178 / 71",
                      "239 / 228 / 141")
rgb2hex <- function(x){rgb(x[1],x[2],x[3], maxColorValue = 255)}

agora.cols.prim <- sapply(strsplit(agora.cols.prim.str," / "), rgb2hex)
agora.cols.1 <- sapply(strsplit(agora.cols.1.str," / "), rgb2hex)
agora.cols.2 <- sapply(strsplit(agora.cols.2.str," / "), rgb2hex)
agora.cols.3 <- sapply(strsplit(agora.cols.3.str," / "), rgb2hex)
```

```{r}
manipulate(
  {ggplot(plotTableF(purchase.tot.bev = purchase.tot.bev, 
                       purchase.tot.diesel = purchase.tot.diesel, 
                       purchase.tot.petrol = purchase.tot.petrol, 
                       deprec.factor = deprec.factor, 
                       ownership.yrs= ownership.yrs, 
                       subsidy.pos = subsidy.pos, 
                       km.yr = km.yr, 
                       e.eur.kwh = e.eur.kwh , 
                       petrol.eur.L = petrol.eur.L,
                       diesel.eur.L = diesel.eur.L, 
                       annual.overhead = annual.overhead), aes(x = engine.type, y = value, fill=cost.component)) +
      scale_fill_manual(values = agora.cols.1) +
      geom_col(position = "stack") + 
      facet_grid(cols=vars(year)) +
      ylab("TCO (Eur)") +
      theme(axis.text.x = element_text(angle = 90))},
  purchase.tot.bev = slider(0,3000,initial = 2000,label = "BEV change in purchase price (Eur/yr)"), 
  purchase.tot.diesel = slider(0,3000,initial = 1000,label = "Diesel change in purchase price (Eur/yr)"), 
  purchase.tot.petrol = slider(0, 3000,initial = 1000,label = "Petrol change in purchase price (Eur/yr)"), 
  deprec.factor = slider(-0.5,0,initial = -0.3 ,label = "Depreciation factor (yrs^(-1))"), 
  ownership.yrs= slider(0,15,initial = 8,label = "Duration of ownership (yrs)"), 
  subsidy.pos = slider(0, 10000,initial = 4000,label = "Point of sale subsidy (Eur)"), 
  km.yr = slider(10000,40000,initial = 15000,label = "Distance driven (km/yr)"), 
  e.eur.kwh = slider(0,1,initial = 0.3,label = "Price of electricity (Eur/kWh)"), 
  petrol.eur.L = slider(0.0,3.0,initial = 1.28,label = "Price of petrol (Eur/L)", step = 0.01), 
  diesel.eur.L = slider(0,3,initial = 1.02,label = "Price of diesel (Eur/L)", step = 0.01), 
  annual.overhead =slider(0,1500,initial = 810,label = "Other annual fixed costs (maintenance, insurance etc.; Eur)")
)

```



